Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2022-01-20 paper/SupMat.Rnw
starting new SI section on methods for quantifying ATA in empirical applicatoins
added empirical applications section

@@

%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decide not to use)
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\olep}{\overline{\epsilon}}

@@

>>>>

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{section}{0}
\renewcommand{\thesection}{S\arabic{section}}

@@

%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decided not to use)
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\olep}{\overline{\epsilon}}
\newcommand{\olE}{\overline{E}}
\newcommand{\olC}{\overline{C}}

@@

\section{Decomposing storage effect in empirical applications}\label{sect:Empircal_app}

The decomposition of storage effects and quantification of contributions of ATAs depends on 
the removal of asymmetric tail associations from the relationship between
coexistence drivers while maintaining their correlation and marginal distributions. 
In the paper section \ref{MT-section:theory}, we describe analytical methods for generating the partial sharp variables, $E_i^\nat$ and $C_{i \bs i}^\nat$, which are distributed in the same way as the defined $E_i$ and $C_{i \bs i}$, but are symmetric in the tails and have a normal copula structure.
The protocol for samples of coexistence drivers in empirical applications is as follows.
Let $\olE_i$ be a sample of a environmental fluctuation and $\olC_{i \bs i}$ be a sample of competitive pressure. Let $\Theta_{\olE}$ and $\Theta_{\olC}$ be standard normal distributions obtained by normalizing the ranks of $\olE_i$ and $\olC_{i \bs i}$, 
such that $\Theta_{\olE} = \frac{rank(\olE_i)}{n_{\olE_i}+1} \circ \varphi^{-1}$ and $\Theta_{\olC} = \frac{rank(\olC_{i \bs i})}{n_{\olC_{i \bs i}}+1} \circ \varphi^{-1}$, where $\varphi$ is the CDF of a standard normal distribution. The normalized correlation of $\olE_i$ and $\olC_{i \bs i}$, denoted by $r$, is the covariance of $\Theta_{\olE}$ and $\Theta_{\olC}$.  
We then simulate an $S$ number of bivariate normal distributions $N_s\left(\vec{0}, \Sigma \right)$
with covariance matrix $\Sigma = \begin{pmatrix} 1&\ddot{\rho} \\ \ddot{\rho}&1\end{pmatrix}$. 
The columns of $(\olE_{i}, \olC_{i \bs i})$ are permutated and saved $S$ amount of times, 
such that the rank of each column matches with the rank of each column of $N_s$. 
To obtain $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$, the median of the permuations is computed. 
Here, $\ddot\rho$ can be chosen a number of ways such that the Pearson, Spearman, or Kendall 
correlation of $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$ matches that of $(\olE_{i}, \olC_{i \bs i})$
or such that $\ddot\rho = r$. The code for this protocol is in \url{"diatom/partialSharp_fxns.R"} and was used in our application to the diatom chemostat system.
